The preceding triplication was slightly contrived because the arrangement of frames at the beginning was so chosen because it satisfied a property as yet unexplained.
This property being that the $R_{ij}$ and $R_{ji}$ occur in the same super-column of the array as frames.
Clearly this must be so in order that permutations, when applied to both or neither of $R_{ij}$ and $R_{ji}$, preserve the contents of the columns as far as is required.

That an arrangement of frames with this property can be guaranteed to exist for any odd integer is fundamental to the generalised theorem.

\begin{lemma}
For all odd $n$ there exists an array with these properties:
\begin{enumerate}
  \item{the entries of the array consist of all the ordered pairs of the set $N = \{1, 2, \ldots, n\}$ once each.}
  \item{the entries of a given row or column contain between them every member of $N$ once as a left member and once as a right member.}
  \item{if $(x,y)$ occurs in a given column of the array $(y,x)$ also occurs in that column.}
\end{enumerate}
\end{lemma}

\begin{proof}
$A_n$ is an $n \times n$ array whose $(i, j)$ entry is the
ordered pair $(j - i + 1, i + j - 1)$ with both elements being
reduced modulo $n$ to lie on the interval $[1, n]$.
\begin{enumerate}
  \item{There are clearly $n^2$ ordered pairs obtainable from
     $N$. $A_n$ has $n^2$ cells, so it is only necessary to
     show that each cell contains a unique pair. For that
     reason consider any two pairs from different cells,
     $(x_1, y_1)$ and $(x_2, y_2)$, for these to be equal
     requires both $x_1 = x_2$ and $y_1 = y_2$.

     Now, $x_1 = x_2$ implies $j_1-i_1 + 1 = j_2-i_2+1$
     While, $y_1 = y_2$ implies $i_1 + j_1 - 1 = i_2 + j_2 - 1$.

     Together these imply,
     \begin{align}
       j_1 - i_1 &= j_2 - i_2 \label{eq:eq1} \\
       i_1 + j_1 &= i_2 + j_2 \label{eq:eq2}
     \end{align}

     \eqref{eq:eq2} gives, $j_2 = i_1 + j_1 - i_2$,
     which on substitution in \eqref{eq:eq1} gives,
     $j_1 - i_1 = i_1 + j_1 - 2i_2$ which implies
     $2i_1 - 2i_2 = 0$.
     
     Therefore $i_1 = i_2$.
     Substituting this into either expression gives
     $j_1 = j_2$.
     Thereby
     contradicting the assumption that the pairs occurred
     in different cells. Hence every cell contains a unique
     pair, so all the ordered pairs from $N$ occur exactly
     once in $A_n$.}
  \item{Consider row $i$ of $A_n$:

    \begin{tabular}{ccccc}
          $j = 1$   &     $j = 2$      & \ldots & $j = - 1$     &     $j = 0$      \\ \hline
       $(2 - i, i)$ & $(3 - i, i + 1)$ & \ldots & $(-i, i - 2)$ & $(1 - i, i - 1)$ 
    \end{tabular}

     Left hand members are $\{2-i, 3-i, \ldots, 1-i\}$, while
     the right hand members are $\{i, i + 1, \ldots, i - 1\}$.
     Both sets
     contain $n$ unique integers on the interval $[1,n]$
     and hence both sets must be $N$, Similarly, consider
     column $j$, the left hand positions are occupied by
     $\{j, j - 1, \ldots, j + 2, j + 1\}$, while the right
     contain $\{j, j + 1, \ldots, j - 2, j - 1\}$.
     For the same reasons both these sets are equal to $N$.}
  \item{Consider a pair $(x,y)$. From the definition of $A_n$,
     $x = j-i + 1$, $y = i + j - 1$. So
     $x + y = j - i + 1 + i + j - 1 = 2j$.
     Therefore $j = \frac{1}{2}(x + y)$. So if $(x, y)$ is in
     column $j$ of $A_n$ then so is $(y, x)$, since
     $\frac{1}{2}(x + y) = \frac{1}{2}(y + x)$.}
\end{enumerate}
\end{proof}

We can now present the general result.

\begin{theorem}
\label{thm:wallis}
If $r$ and $n$ are odd integers such that $r \geq n$, and if there is a Room square $R$ of side $r$, then there is a Room square of side $rn$.
\end{theorem}

\begin{proof}
Let $r = 2d + 1$ and $n = 2t + 1$.

For a given $i$ select $n$ permutations as follows:

\begin{enumerate}
  \item{$\phi _{jk} = \phi _{jl}$ if, and only if, $(k, l)$, and $(l, k)$ appear in column $j$ of $A_n$.}
  \item{If cell $(j, j)$ of $A_n$ contains $(x, y)$ then $\phi _{jx} = \phi _{jy} = id$ (the identity permutation).}
  \item{All the $\phi _{jk} (\neq id)$ are selected from the permutations associated with $R$, according to Lemma \ref{lem:permute}.}
\end{enumerate}

Now a Room square of side $rn$ is constructed by replacing each entry $(k, l)$ of $A_n$ by $R_{kl} \phi _{jk}$ (the array $R_{kl}$ under the column permutation $\phi _{jk}$), where $(k, l)$ is in column $j$ of $A_n$.

The resulting array has each element of
\begin{equation}
S = \{0_1,1_1,...,(r-1)_1,0_2,1_2,...,(r-1)_2,0_n,1_n,...,(r-1)_n\}
\end{equation}
appearing exactly once in each row and column, except that $x_j$ is missing from row and column
$x_j,$ $1 \leq j \leq n 0 \leq x \leq (r-1)$, and $x_k$ and $x_l$ are missing from column $(x \phi _{jk})_j$ for each entry $(k,l)$ in column $j$ of $A_n$.

The array also contains every unordered pair from $S$ exactly once, except those of the form $\{x_k,x_l\}$.
Now, for each $k$, if $(k, l)$ is an entry of column $j$ of $A_n$ put $\{x_k, x_l\}$ in $(x_j, (x \phi _{jk})_j)$, using $\{\infty, x_j\}$ instead of $\{x_j, x_j\}$ in every case.

The completed array contains each of $\{\infty\} \cup S$ exactly once per row and column and every unordered pair from the same set exactly once.

Finally, map $\{\infty\} \cup S$ onto $\{\infty \} \cup Z_{rn}$ by replacing every $x_i$ by $x + r(i - 1)$.
The finished array is a Room square of side $rn$.
\end{proof}

\begin{example}
Looking back at the previous example of triplication.

\begin{equation}
 A = \begin{bmatrix}
  1,1 & 2,2 & 3,3 \\
  3,2 & 1,3 & 2,1 \\
  2,3 & 3,1 & 1,2 \\
 \end{bmatrix}
\end{equation}

Now, in column 1 $(3, 2)$ and $(2, 3)$ both appear, so $\phi _{13} = \phi _{12}$.
While in column 2 occur both (1,3) and (3,1), so $\phi _{21} = \phi _{23}$.
Furthermore, both $(2, 1)$ and $(1, 2)$ appear in the third column, so $\phi _{32} = \phi _{31}$.
The diagonal pairs are $(1, 1)$, $(1, 3)$, $(1, 2)$ so
\begin{equation}
\phi _{11} = \phi _{21} = \phi _{23} = \phi _{31} = \phi _{32} = id
\end{equation}

The remaining permutations, $\phi _{12}, \phi _{13}, \phi _{22}, \phi _{33}$ are chosen according to the Lemma \ref{lem:permute}
and the following array is the Room square in \ref{eq:roomtwentyone} once the missing pairs have been placed and the transformation to $\{\infty, 0, 1, ..., 20\}$ made.

\begin{equation}
  \begin{bmatrix}
   R_{11}\phi_{11} & R_{22}\phi_{22} & R_{33}\phi_{33} \\
   R_{32}\phi_{13} & R_{13}\phi_{21} & R_{21}\phi_{32} \\
   R_{23}\phi_{12} & R_{31}\phi_{23} & R_{12}\phi_{31} \\
  \end{bmatrix}
  =
  \begin{bmatrix}
      R_{11}id    & R_{22}\phi_{2} & R_{33}\phi_{3} \\
   R_{32}\phi_{1} &   R_{13}id     &    R_{21}id    \\
   R_{23}\phi_{1} &   R_{31}id     &    R_{12}id    \\
  \end{bmatrix}
\end{equation}

\end{example}
